The impact of gas bulk rotation on the Lyman-$\alpha$ line

### 2014-01-27 paper/ms.tex
editando

\begin{figure}
    \includegraphics[width=0.45\textwidth]{NewWidthVvsVmax.png}
    \caption{Half-width for the non-dusty models as a function of
      rotational velocity $V_{\rm max}$. Continuous (dashed) lines
      correspond to central (homogeneous) source
      distributions. \label{fig:widthvsvelocity}} 
\end{figure}

@@

\begin{figure*}
\begin{center}
  \includegraphics[width=0.45\textwidth]{f5_1.eps}
  \includegraphics[width=0.45\textwidth]{f5_2.eps}\\
  \includegraphics[width=0.45\textwidth]{f5_3.eps}
  \includegraphics[width=0.45\textwidth]{f5_4.eps}    
\end{center}
    \caption{2D histogram of $N_{\rm scatt}$ vs $x$. The upper (lower) panels
      show the homogeneous (central) source distribution. Left
      corresponds to the static case and right
      $V_{\rm max}=300km/s$. The color scale is logarithmic on the
      number of photons with given values of $N_{\rm scatt}$ and
      $x$. \label{fig:Nscatt2D}}   
\end{figure*}

The number of scatterings affects the escape frequency of a \ly photon.
In static cases, a large value of the optical depth correlates with a
high number of scatterings, increasing the probability of finding a
\ly photon far from the line's center. The result is that he peak maxima
shifts from the center as the amount of neutral hydrogen
increases. This can be precisely quantifyied in for the static slab
with central sources, in that case the maxima position are related to
the optical depth and the temperature as $x_{m}=\pm 1.066(a\tau_{\rm
  H})^{1/3}$ \citep{Harrington73}.

%%%% JEFR.

In Figure \ref{fig:Nscatt} we show the average number of scatterings
$\langle N_{\rm scatt}\rangle$ as a function of the rotational velocity
$V_{\rm max}$. For the central distributions the average number of
scatterings has a modest change for increasing rotational
velocities, $\langle N_{\rm scatt}\rangle$ changes less than $0.5\%$
for different velocities. In the same set-up the average
number of scatterings is proportional to the optical depth, as
expected in analogy from the analytic result for the central
infinite-slab $\langle N_{\rm  scatt}\rangle=1.612\tau_{\rm   H}$
\citep{Adams72,Harrington73}. In the spheres with centrally
distributed sources we find $\langle N_{\rm
  scatt}\rangle= (1.50, 1.00, 0.92)\tau_{\rm   H}$ for optical depths
$\tau_{\rm H} = (10^{5}, 10^{6}, 10^{7})$

Figure \ref{fig:Nscatt} shows that for the homogeneous distribution
there is a clear decrease of $\langle N_{\rm  scatt}\rangle$ as the
$V_{\rm max}$ increases. This effect is more pronounced for lower
values of the optical depth. For $\tau_{\rm H}=10^5$ the average
number of scatterings decreases by $61\%$ at $V_{\rm max}=300\kms$ in
comparison to the static case.

The analytic expectation for the average number of scatterigs in a
slab with homogeneously distributed sources is $\langle N_{\rm
  scatt}\rangle=1.16\tau_{\rm   H}$ \citep{Harrington73}. This is
$0.72$ times lower than the same case with photons emitted in the
slab's center. In our case we find that for the static case and the
homogeneously distributed sources $\langle N_{\rm scatt}\rangle=
(0.99, 0.59, 0.51)\tau_{\rm   H}$, this represents factors of $(0.66,
0.59, 0.55)$ lower than the centrally emitted  photons.

In order to gain a deeper understanding of these results we make 2D
histograms for the number of scatterings as a function of the outgoing
dimensionless frequency $x$. In Figure \ref{fig:Nscatt2D} we show
such histogram in the case $\tau_{\rm H}=10^5$ for the
static case and $V_{\rm max}=300$\kms. The upper (lower) panels show the
results for the homogeneous (central) source distribution. The color
scale is logarithmic in the number of photons at a certain value
$x-N_{\rm scatt}$.

The top right panel of Figure \ref{fig:Nscatt2D} supports our hypothesis about
the photo-sphere in the homogeneous distribution. In that case most of
the photons that left with $x\sim 0$ have escaped with less than $10$
scatterings. This explains the origin of a single central
peak. However, for a central distribution the situation is
different (lower panels). In this case the number of scatterings remains high, in the
order of the optical depth, but the two peaks do get closer to each
other. In this case the physical picture is that each scattering, due
to the bulk velocity of the gas, is inefficient in driving the photon
outside the line center.

@@

\begin{figure}
  \includegraphics[width=0.45\textwidth]{f6.eps}
   \caption{Escape fraction as a function of rotational velocity, for these models we took $\tau_{a}=1$. The
     continuous (dashed) lines correspond to central (homogeneous)
     models.
     \label{fig:efvsv}}
\end{figure}

\begin{figure}
  \includegraphics[width=0.45\textwidth]{f7.eps}
 \caption{Escape fraction as a function of the
   product $(a\tau_{\rm H})^{1/3}\tau_{a}$. The analytic solution for
   the infinite slab is shown as a continuous line. Different dashed
   lines correspond to different rotational velocities, for this models 
   $\tau_a = 1$.
   \label{fig:efvsNeufeld}}   
\end{figure}

The average number of scatterings is also reflected in the amount of
photons absorbed by dust. Under these considerations we do not expect
any change in the escape fraction for a dusty cloud with central
source of radiation given that the number of scatterings remains close
to constant. On the other hand, in the case of a homogeneous
radiation source the number of scatterings drops as $V_{\rm max}$
increases, which has to be reflected as an increasing escape fraction.

Figure \ref{fig:efvsv} shows the dependence of the escape fraction as
a function of the maximum rotational velocity, confirming our expectations. For the central source
distribution the escape fraction barely shows any change, while for
the homogeneous case there is a clear increase in the escape fraction for
high rotational velocities.  Rotation has a higher relative impact in
the models with low optical depth $\tau_{\rm H}=10^5$,$10^{6}$, where
it can rise from $(0.26, 0.28)$ ($V_{\rm}=0$\kms) up to $(0.37, 0.39)$ ($V_{\rm
  rot}=300$ \kms).

In Figure \ref{fig:efvsNeufeld} we put these results in the context of
the analytic solution for the infinite slab \citep{Neufeld90}. In
Neufeld's set-up the analytic solution depends uniquely on the product
$(a\tau_{\rm   H})^{1/3}\tau_{a}$, an approximation that is valid only
in the limits where $a\tau_{\rm   H}\gg 1$. The dashed lines in Figure
\ref{fig:efvsNeufeld} correspond to the cases of different
velocities. First of all we note that the escape fraction does not
increase from $\tau_{\rm H}=10^{5}$ to $\tau_{\rm H}=10^{6}$, this is
explained since there is a transition from a opaque to an extremely opaque medium
which affect affects how \ly photons escape from the medium: single flight 
vs single excursion.
 %we are in a regime where the condition for the analytic expectations  ($a\tau_{h}\gg 1$) does not hold. 
This figure also shows that the escape fraction is higher by factors of $2-10$
than the expectation for a slab with centrally distributed
sources. Furthermore, the decrease of the escape fraction with
increasing optical depth is also slower.

\subsection{Off-Centered emission}
\label{sec:off-center}
\begin{figure}
  \includegraphics[scale=0.3]{f8.eps}
  \caption{The two solid circles show the two different spherical
    off-centered emitting  regions. The angular velocity vector is
    defined to be along the $\hat{\i}$ direction (where the reader is) 
    and the observer is along the $\hat{\j}$ direction. 
    \label{fig:OCspheres}} 
\end{figure}

>>>>

\begin{figure}
    \includegraphics[width=0.45\textwidth]{NewWidthVvsVmax.eps}
    \caption{Half-width for the non-dusty models as a function of
      rotational velocity $V_{\rm max}$. Continuous (dashed) lines
      correspond to central (homogeneous) source
      distributions. \label{fig:widthvsvelocity}} 
\end{figure}

@@

\begin{figure*}
\begin{center}
  \includegraphics[width=0.45\textwidth]{f5_1.eps}
  \includegraphics[width=0.45\textwidth]{f5_2.eps}
  \includegraphics[width=0.45\textwidth]{f5_3.eps}
  \includegraphics[width=0.45\textwidth]{f5_4.eps}    
\end{center}
    \caption{2D histogram of $N_{\rm scatt}$ vs $x$. The upper (lower) panels
      show the homogeneous (central) source distribution. Left
      corresponds to the static case and right
      $V_{\rm max}=300\kms$. The color scale is logarithmic on the
      number of photons with given values of $N_{\rm scatt}$ and
      $x$. \label{fig:Nscatt2D}}   
\end{figure*}

The number of scatterings affects the escape frequency of a \ly photon.
In static cases, a large value of the optical depth correlates with a
high number of scatterings, increasing the probability of finding a
\ly photon far from the line's center. As a result the peak maxima
shifts from the center as the amount of neutral hydrogen
increases. This can be precisely quantifyied in the static slab
with central sources where the maxima position are related to
the optical depth and the temperature as $x_{m}=\pm 1.066(a\tau_{\rm
  H})^{1/3}$ \citep{Harrington73}. In the same model the average
number of scatterings depends only on the optical depth $\langle
N_{\rm  scatt}\rangle=1.612\tau_{\rm   H}$
\citep{Adams72,Harrington73}, in the case of homogeneously distributed
sources $\langle N_{\rm   scatt}\rangle=1.16\tau_{\rm   H}$
\citep{Harrington73}.

In Figure \ref{fig:Nscatt} we show the average number of scatterings
$\langle N_{\rm scatt}\rangle$ as a function of the rotational velocity
$V_{\rm max}$. For the central source distribution the average number of
scatterings $\langle N_{\rm   scatt}\rangle$ changes less than $0.5\%$
for different velocities. In this setup the number of scatterings is
proportional to the optical depth, with $\langle N_{\rm
  scatt}\rangle= (1.50, 1.00, 0.92)\tau_{\rm   H}$ for optical depth
values of $\tau_{\rm H} = (10^{5}, 10^{6}, 10^{7})$, respectively

From Figure \ref{fig:Nscatt} is also clear that for the homogeneous
distribution there is a clear decrease of $\langle N_{\rm
  scatt}\rangle$ as the $V_{\rm max}$ increases. For instance, for
$\tau_{\rm H}=10^5$ the average number of scatterings decreases by
$61\%$ at $V_{\rm max}=300\kms$ in comparison to the static case.

In this case we find that for a static sphere $\langle N_{\rm
  scatt}\rangle= (0.99, 0.59, 0.51)\tau_{\rm   H}$, this
representsfactors of $(0.66, 0.59, 0.55)$ lower than the centrally
emitted photons. In contrast the analytic solution for the homogeneous
slab gives $0.72$ times less scatterings in the homogeneous case than
in the central one.

In order to gain a deeper understanding of these results and explain
the results for the maxima in the line morphology we make 2D
histograms for the number of scatterings as a function of the outgoing
dimensionless frequency $x$. In Figure \ref{fig:Nscatt2D} we show
such histogram in the case $\tau_{\rm H}=10^5$ for the
static case and $V_{\rm max}=300$\kms. The upper (lower) panels show the
results for the homogeneous (central) source distribution. The color
scale is logarithmic in the number of photons at a certain value
$x$-$N_{\rm scatt}$.

The top right panel of Figure \ref{fig:Nscatt2D} (homogeneous sources,
high rotational velocity) supports our hypothesis about
the photo-sphere in the homogeneous distribution. In this case most of
the photons that left with $x\sim 0$ have escaped with less than $10$
scatterings. This also explains the decrease in the average number of
scatterings observed in Figure \ref{fig:Nscatt}.

However, for a central distribution the situation is quite different
(lower panels). In this case the number of scatterings remains high,
in the order of the optical depth, but the two peaks do get closer to each
other. Here the most probable physical picture is that each
scattering, due to the bulk velocity of the gas, is inefficient in
driving the photon outside the line center.

@@

\begin{figure}
  \includegraphics[width=0.45\textwidth]{f6.eps}
  \caption{Escape fraction as a function of rotational velocity. All
    these models have $\tau_{a}=1$. The continuous (dashed) lines
    correspond to central (homogeneous) models.
    \label{fig:efvsv}}
\end{figure}

\begin{figure}
  \includegraphics[width=0.45\textwidth]{f7.eps}
  \caption{Escape fraction as a function of the
    product $(a\tau_{\rm H})^{1/3}\tau_{a}$. The analytic solution for
    the infinite slab is shown as a continuous line. Different dashed
    lines correspond to different rotational velocities.
    {\bf IS THE ALBEDO TAKEN INTO ACCOUNT?}
    \label{fig:efvsNeufeld}}   
\end{figure}

In the case of dusty clouds, all changes in the average number
of scatterings presented in the previous subsection have a direct
bearing on the escape fraction. From the results we have just obtained
we can predict that for central source distributions the escape
fraction will barely change, while larger effects should be expecte
for homogeneously distributed sources.

Figure \ref{fig:efvsv} confirm these expectations. It shows the escape
fraction as a function of the rotational velocity. The curves for the
central source distribution stay flat, while for the homogeneous case
there is a clear rise with rotational velocity.  Rotation has a higher relative impact in
the models with low optical depth. For instance, in models with
$\tau_{\rm H}=10^5$, the static scape fraction is $0.26$ and increases
to $0.37$ for $V_{\rm max}=300$ \kms. Table \ref{table:escape} lists
all the values for the escape fraction {\bf We also need a table
  listing the escape  fraction values.}

In Figure \ref{fig:efvsNeufeld} we put these results in the context of
the analytic solution for the infinite slab \citep{Neufeld90}. In
Neufeld's set-up the analytic solution depends uniquely on the product
$(a\tau_{\rm   H})^{1/3}\tau_{a}$, an approximation that is valid only
in the limit $a\tau_{\rm   H}\gg 1$. The dashed lines in Figure
\ref{fig:efvsNeufeld} show the results for the different rotational
velocities. First of all we note that the escape fraction does not
increase from $\tau_{\rm H}=10^{5}$ to $\tau_{\rm H}=10^{6}$, a fact
that is explained since there is a transition from a opaque to an extremely opaque medium
which affect affects how \ly photons escape from the medium: single flight 
vs. single excursion. {\bf Que quiere decir eso de single flight y
  single excursion, porque explica eso la fraccion de escape?} 
 %we are in a regime where the condition for the analytic expectations
 %($a\tau_{h}\gg 1$) does not hold.

\subsection{Anisotropy and off-centered emission}

\label{sec:off-center}
\begin{figure}
  \includegraphics[scale=0.3]{f8.eps}
  \caption{The two small circles show the two different spherical
    off-centered emitting regions with respect to the larger spherical
    gas distribution. The angular velocity vector is
    defined to be along the $\hat{\i}$ direction (outwards the
    page). In order to study a possible anysotropy we place an
    observer along the the $\hat{\j}$ direction. The vector
    $\hat{k}_{\rm out}$ represents the outgoing direction of an
    escaping \ly photon. {\bf Si el observador esta en j, entonces las
    dos esferas no deberian estar sobre el eje k? Lo mejor es ademas
    hacer las dos circulos pequenos punteados}
    \label{fig:OCspheres}} 
\end{figure}

Rotation breaks the spherical symmetry of the static model and
introduces a prefered axis
